Using Probabilistic Knockoffs of Binary Variables to Control the False Discovery Rate

### 2015-07-26 masters_paper/paper.tex
more paper

\newcommand{\deuc}{d_{\mathrm euc}}
\newcommand{\dtaxi}{d_{\mathrm taxi}}
\newcommand{\ddisc}{d_{\mathrm disc}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

@@

    \[ \beta(\lambda) = \textrm{argmin}_\mathbf b \left\{\frac{1}{2}\|\mathbf y - X_L\mathbf b\|^2_2 + \lambda\|b\|_1 \right\}\]

@@

\section{Issues With Deterministic Knockoffs} 
    The knockoff filter and $W$ statistics have a very natural extension to generalized linear models. Here, where $l(\beta\st X, \tilde X)$ is the log likelihood of coefficients $\beta$ given the model, the $L1$ regularized regression model will have estimated coefficient vector
    \[ \beta(\lambda) = \textrm{argmin}_\mathbf b \left\{l(\beta\st X, \tilde X) + \lambda\|b\|_1 \right\}\]
    The $Z_j$ and $\tilde Z_j$ are then, once again, the largest $\lambda$ such that the given original or knockoff feature has a positive coefficient, and then $W_j$ is constructed exactly as before. The problem with this is that if the given GLM doesn't have the same sufficient statistics as linear regression, there is no theoretical guarantee that the knockoff filter will control FDR. This is since, for null $X_j$, there is no theory to guarantee the $Z_j$ and $\tilde Z_j$ are exchangeable, and thus $W_j$ is not guaranteed to be symmetrically distributed around $0$. \par
    \begin{figure}[h]
        \begin{center}
        \includegraphics[width=14cm]{images/entryrate_original_logit}
    \end{center}
        \caption{The mean cumulative portion of $W$ statistic which are positive by absolute rank of the value. The mean is over 1000 simulations. On the left side, the features are drawn from a random standard normal distribution. On the right side, they are drawn as Bernoulli with probability $.25$ of being 1. The top has 50 features uncorrelated with the outcome, while the bottom has 40 features uncorrelated with the outcome and 10 correlated with it.}
    \end{figure}

@@

            \[\mathbf a_i^{(k+1)} = \left[H f\left(\mathbf a_i^{(k)}\right)\right]^{-1}\left[f\left(\mathbf a_i^{(k)}\right)-\mathbf m_i\right] \]
            To do so, the Jacobian matrix is calculated as 
        \[ J\left( f\left(\mathbf a_i\right)\right) = \sum_{\mathbf x_{-i} \in \{0,1\}^{i-1}} \frac{1}{c}L \left(\mathbf x_{-i}^TA_{-i}\mathbf x_{-i}\right) L'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf x_{-i}^T & 1 \end{array}\right] \]

@@

            \[J\left( f\left(\mathbf a_i\right)\right) \approx \frac{1}{K}\sum_{k=1}^K L'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf [x_{-i}^{(k)}]^T & 1 \end{array}\right] \]

@@

            \item As $p$ grows large, and $K$ necessarily along with it, the cost of calculating the Jacobian matrix grows far quicker than the cost of calculating $f(\mathbf a)$. This is since it has $O(Kp^2)$ complexity compared to $O(Kp)$ complexity for $f(\mathbf a)$. This can quickly become prohibitive. One might instead estimate the full Jacobian based on first differences, but in practice Anderson Mixing, a quasi-Newtonian method, works well. In general, quasi-Newtonian methods perform Newton-Raphson iterations, but replace the full inverted Jacobian with an approximation. {\bf MORE DETAILS} 

@@

        Random binary knockoffs which produce the desired cross moment matrix $M_L$ need not arise from a generative model for $\mathbf{\tilde x}$ however. It is sufficient to have a random matrix $\tilde X \in \{0,1\}^{n\times p}$ such that 
        \[ \E_{\tilde X}(X^T \tilde X) = \Sigma - \diag{s} \quad \& \quad \E_{\tilde X}(\tilde X^T \tilde X) = \Sigma\]
        The simplest way to do this is based on a probability matrix $P\in (0,1)^{n\times p}$ where each $X_{ij}$ is drawn independently as $X_{ij}\sim\textrm{Bernoulli}(P_{ij})$. This requirement can be stated as the quadratic feasibility problem
            \begin{center}
                \begin{tabular}{r l}
                    exists     & $P$ \\
                    subject to & $X^TP=\Sigma-\diag{s} $ \\
                               & $P^TP=\Sigma $ \\
                               & $0 \leq P \leq 1$
                \end{tabular} 
            \end{center}
        If such a matrix exists and can be found, then $\tilde X$ can easily be generated. However, the constraints likely result in infeasibility much of the time. Instead, one can try to minimize how far the expected moments given $P$ would be from those desired for $M_L$. Ideally, this would take the form
            \begin{center}
                \begin{tabular}{r l}
                    minimize     & $\|X^TP-(\Sigma-\diag{s})\|_{fro}^2 + \|P^TP-\Sigma\|_{fro}^2 $\\
                    subject to   & $ \mathbf 1^T P = \mathbf m_L $ \\
                                 & $0 \leq P \leq 1$
                \end{tabular} 
            \end{center}
            Here, the marginal distribution of $\tilde X_i$ will be the same as $X_i$, and the crossmoments shouldn't deviate to far from $M_L$. This problem can be stated as the quadratic program via the introduction of slack variables:
            \begin{center}
                \begin{tabular}{r l}
                    minimize     & $\|W\|_{fro}^2 + \|V\|_{fro}^2 $ \\
                    subject to   & $ X^TP-(\Sigma-\diag{s})\leq W $ \\
                                 & $ P^TP-\Sigma \leq V $ \\
                                 & $ \mathbf 1^T P = \mathbf m_L $ \\
                                 & $0 \leq P \leq 1$
                \end{tabular} 
            \end{center}
            Though there are nice tools and theory for solving quadratic programs, $n\times p$ would have to be fairly small for this optimization problem to be computationally reasonable in practice. Ideally, there would be a method to breakdown this problem into a series of smaller optimization problems which would yield the same solution. A optimistic numerical analyst would hope that iterating through the columns, optimizing each in individually, would achieve this.

>>>>

\newcommand{\deuc}{d_{\mathrm euc}}
\newcommand{\dtaxi}{d_{\mathrm taxi}}
\newcommand{\ddisc}{d_{\mathrm disc}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\DeclareMathOperator*{\argmin}{arg\min}

@@

    \[ \beta(\lambda) = \argmin_\mathbf b \left\{\frac{1}{2}\|\mathbf y - X_L\mathbf b\|^2_2 + \lambda\|b\|_1 \right\}\]

@@

\section{Issues With Deterministic Knockoffs} 
    The knockoff filter and $W$ statistics have a very natural extension to generalized linear models. Here, where $l(\beta\st X, \tilde X)$ is the log likelihood of coefficients $\beta$ given the model, the $L1$ regularized regression model will have estimated coefficient vector
    \[ \beta(\lambda) = \argmin_\mathbf b \left\{l(\beta\st X, \tilde X) + \lambda\|b\|_1 \right\}\]
    The $Z_j$ and $\tilde Z_j$ are then, once again, the largest $\lambda$ such that the given original or knockoff feature has a positive coefficient, and then $W_j$ is constructed exactly as before. The problem with this is that if the given GLM doesn't have the same sufficient statistics as linear regression, there is no theoretical guarantee that the knockoff filter will control FDR. This is since, for null $X_j$, there is no theory to guarantee the $Z_j$ and $\tilde Z_j$ are exchangeable, and thus $W_j$ is not guaranteed to be symmetrically distributed around $0$. \par
    \begin{figure}[h]
        \begin{center}
        \includegraphics[width=14cm]{images/entryrate_original_logit}
    \end{center}
        \caption{The mean cumulative portion of $W$ statistic which are positive by absolute rank of the value. The mean is over 1000 simulations. On the left side, the features are drawn from a random standard normal distribution. On the right side, they are drawn as Bernoulli with probability $.25$ of being 1. The top has 50 features uncorrelated with the outcome, while the bottom has 40 features uncorrelated with the outcome and 10 correlated with it.}
    \end{figure}

@@

            \[\mathbf a_i^{(k+1)} = \mathbf a_i^{(k)} - \left[J\left(\mathbf a_i^{(k)}\right) \right]^{-1}\left[f\left(\mathbf a_i^{(k)}\right)-\mathbf m_i\right] \]
            where $J$ is the Jacobian of $f$. To do so, the Jacobian matrix is calculated as 
        \[ J\left( \mathbf a_i\right) = \sum_{\mathbf x_{-i} \in \{0,1\}^{i-1}} \frac{1}{c}L \left(\mathbf x_{-i}^TA_{-i}\mathbf x_{-i}\right) L'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf x_{-i}^T & 1 \end{array}\right] \]

@@

            \[J\left( \mathbf a_i\right) \approx \frac{1}{K}\sum_{k=1}^K L'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_{-i}^{(k)} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf [x_{-i}^{(k)}]^T & 1 \end{array}\right] \]

@@

            \item As $p$ grows large, and $K$ necessarily along with it, the cost of calculating the Jacobian matrix grows far quicker than the cost of calculating $f(\mathbf a)$. This is since it has $O(Kp^2)$ complexity compared to $O(Kp)$ complexity for $f(\mathbf a)$. This can quickly become prohibitive. One might instead estimate the full Jacobian based on first differences, but in practice Anderson Mixing, a quasi-Newtonian method, works well. In general, quasi-Newtonian methods perform Newton-Raphson iterations but replace the full inverted Jacobian at each itteration $k$ with an approximation $G_k$. Thus, to solve for a variable $\mathbf z$ such that $f(\mathbf z)=\mathbf 0$, the best approximation for the zero is updated as $\mathbf z_{k+1}=\mathbf z_k - G_k f(\mathbf{z}_k)$. \par
    Anderson mixing does this implicitly, with no extra recalculation of $f$, by approximating $f$ with the affine space spanning the previous $m$ evaluations of $f$, at $\mathbf z_{k}$ through $\mathbf z_{k-m+1}$. The best approximation for $f(\mathbf z)=0$ in this space is defined by
    \[ \mathbf \alpha = \argmin_{\|\mathbf w\|_1=1} \|\sum_{i=1}^m f(\mathbf z_{k-i+1}) w_i \|_2 \]
    This leads to the updated approximation
    \[ \mathbf z_{k+1} = \sum_{i=1}^{m} w_i \mathbf z_{k-i+1} \]
    There is no explicit calculation of $G_k$ needed here, and to do so in practice would be slower. However, it's presence can be seen through an equivalent formulation. Let 
    \[ \mathcal F = [(f(\mathbf z_k) - f(\mathbf z_{k-1})) \ldots (f(\mathbf z_k) - f(\mathbf z_{k-m}))] \]
    \[ \mathcal Z = [(\mathbf z_k - \mathbf z_{k-1})\ldots (\mathbf z_k - \mathbf z_{k-m})] \]
    This means $\mathbf \alpha$ is now
    \[ \mathbf \alpha \; = \;\argmin_{\|\mathbf w\|_1=1} \|f(\mathbf z_{k}) - \mathcal  F \mathbf w\|_2\; = \;(\mathcal F^T \mathcal F)^{-1}\mathcal F^T\mathbf z_k \]
    so 
    \[ \mathbf z_{k+1} \;= \;\mathbf z_{k+1} - \mathcal Z \alpha\; = \;z_{k+1} + \mathcal Z(\mathcal F^T \mathcal F)^{-1}\mathcal F^T\mathbf z_k \]
    Which yields the approximated inverse Jacobian $G_k= \mathcal Z(\mathcal F^T \mathcal F)^{-1}\mathcal F^T$

@@

        Random binary knockoffs which produce the desired cross moment matrix $M_L$ need not arise from a generative model for $\mathbf{\tilde x}$ however. It is sufficient to have a random matrix $\tilde X \in \{0,1\}^{n\times p}$ such that 
        \[ \E_{\tilde X}(X^T \tilde X) = \Sigma - \diag{s} \quad \& \quad \E_{\tilde X}(\tilde X^T \tilde X) = \Sigma\]
        The simplest way to do this is based on a probability matrix $P\in (0,1)^{n\times p}$ where each $X_{ij}$ is drawn independently as $X_{ij}\sim\textrm{Bernoulli}(P_{ij})$. This requirement can be stated as the quadratic feasibility problem
                    exists     & $P$ \\
                    subject to & $X^TP=\Sigma-\diag{s} $ \\
                               & $P^TP=\Sigma $ \\
                               & $0 \leq P \leq 1$
        If such a matrix exists and can be found, then $\tilde X$ can easily be generated. However, the constraints likely result in infeasibility much of the time. Instead, one can try to minimize how far the expected moments given $P$ would be from those desired for $M_L$. Ideally, this would take the form
                    minimize     & $\|X^TP-(\Sigma-\diag{s})\|_{fro}^2 + \|P^TP-\Sigma\|_{fro}^2 $\\
                    subject to   & $ \mathbf 1^T P = \mathbf m_L $ \\
                                 & $0 \leq P \leq 1$
            Here, the marginal distribution of $\tilde X_i$ will be the same as $X_i$, and the crossmoments shouldn't deviate to far from $M_L$. This problem can be stated as the quadratic program via the introduction of slack variables:
                    minimize     & $\|W\|_{fro}^2 + \|V\|_{fro}^2 $ \\
                    subject to   & $ -W \leq X^TP-(\Sigma-\diag{s})\leq W $ \\
                                 & $ -V \leq P^TP-\Sigma \leq V $ \\
                                 & $ \mathbf 1^T P = \mathbf m_L $ \\
                                 & $0 \leq P \leq 1$
            Though there are nice tools and theory for solving quadratic programs, $n\times p$ would have to be fairly small for this optimization problem to be computationally reasonable in practice. Ideally, there would be a method to breakdown this problem into a series of smaller optimization problems which would yield the same solution. An optimist might hope that iterating through the columns of $P$, optimizing each to hit its desired crossmoments individually conditioned on the estimate of the rest of $P$, would achieve this. In practice, this does seem to yield reasonable results, and is far quicker than other methods for large $p$.In particular, this method seemed to work well:
        \item Optimize initial $P_i$ as such that $X_j^T P_i = M_ij$ if $i\neq j$ and $X_i^T P_i = \sigma_i(1-s)$.   
                    minimize     & $ \left(\frac{1}{n}X_i^T P_i - (M_{ij}-\sigma_i(1-s))\right)^2 + \sum_{i\neq j}(\frac{1}{n}X_j^T P_i - M_ij)^2$ \\
                    subject to   & $ \frac{1}{n}\mathbf 1^T P_i = M_{ii} $ \\
                                 & $0 \leq P_i \leq 1$
        \item Draw initial $\tilde X_i$ as independent Bernoulli with probability $P_i$.
        \item Iterate through $1\leq i \leq p$, fitting new $P_i$ such that 
                    minimize     & $ \left(\frac{1}{n}X_i^T P_i - (M_{ij}-\sigma_i(1-s))\right)^2 + \sum_{i\neq j}(\frac{1}{n}X_j^T P_i - M_ij)^2 + \sum_{i\neq j}(\frac{1}{n}\tilde X_j^T P_i - M_ij)^2$  \\
                    subject to   & $ \frac{1}{n}\mathbf 1^T P_i = M_{ii} $ \\
                                 & $0 \leq P_i \leq 1$
